|

### 2017-11-24 boettiger-eddelbuettel.Rmd

The `debian:testing` stack aims to make the most efficient use of
upstream builds: the pre-compiled `.deb` binaries provided by the
Debian repositories.  It is both quicker and easier to install software
from binaries, since the package manager (`apt`) manages the necessary
(binary) dependencies and bypasses the time-consuming process of compiling
from source.  Basing this stack on `debian:testing` means that much more
recent versions of commonly-used libraries and compilers are available
as binaries than would be found in a Debian stable release.  In order
to provide access to the most recent available binaries, this stack uses
apt-pinning \citep{apt_pinning} to allow the `apt` package manager to also
install binaries from `debian:unstable`, which represents the most recent,
bleeding edge of packages built for Debian when necessary.  For instance,
the `r-base` image provided by Rocker installs the most recent version of
R as a binary from Debian `unstable`.  Similarly, recent versions of many
popular R packages can also be installed through the package manager,
_e.g._, `apt-get install r-cran-xml`.  This can be particularly helpful
for packages with external system dependencies (such as `libxml2-dev`
in this example) which cannot be installed from the R console as they 
are system dependencies rather than R packages installed from within R.
We should note, however, that only about 500 of the over
11,000 CRAN packages are available as Debian packages.

As the names `testing` and `unstable` imply, this approach is not
without challenges.  The particular version of any given package can
change as packages move from `unstable` into `testing`. New versions
are sent to `unstable` during the normal course of Debian development.
This can occasionally break an previously-working installation command
in a Dockerfile until the maintainer redirects the package manager to
install a package from the `unstable` sources that could previously
be installed from `testing`, or vice versa (using the `-t` option in
`apt`).  That said, packages only migrate from 
`unstable` to `testing` after a period of several days---and if the 
migration and installation of the particular version is free of 
interactions with other packages in their dependency graph. That way,
`unstable` serves as validation lab which leaves `testing` reasonably 
stable yet current.

Relative to `stable`, the `testing` stack thus offers the following advantages:

1. These Dockerfiles are easy to develop and extend because almost all software can be installed through the package manager.
2. These Dockerfiles always install the most recent available software.
3. These Dockerfiles can almost always build relatively quickly.

and these down-sides:

1. These Dockerfiles require occasional minor maintenance to ensure successful builds. (_e.g._, changing an installation directive from `unstable` to `testing` or vice versa).
2. The resulting images are inherently dynamic: rebuilding the same Dockerfile months or years apart will generate images with significantly different versions of software installed.
3. The use of apt pinning may be unfamiliar to some users, where the user must ultimately be responsible to ensure compatibility \citep{apt_pinning}.

>>>>

The `debian:testing` stack aims to make the most efficient use of
upstream builds: the pre-compiled `.deb` binaries provided by the
Debian repositories.  It is both quicker and easier to install software
from binaries, since the package manager (`apt`) manages the necessary
(binary) dependencies and bypasses the time-consuming process of compiling
from source.  Basing this stack on `debian:testing` means that much more
recent versions of commonly-used libraries and compilers are available
as binaries than would be found in a Debian stable release.  In order
to provide _optional_ access to the most recent available binaries, this stack uses
apt-pinning \citep{apt_pinning} to allow the `apt` package manager to _selectively_
install binaries from `debian:unstable`, which represents the most recent
set of packages built for Debian. Similarly, recent versions of many
popular R packages can also be installed pre-built through the package manager,
_e.g._, `apt-get install r-cran-xml`.  This can be particularly helpful
for packages with external system dependencies (such as `libxml2-dev`
in this example) which cannot be installed from the R console as they 
are system dependencies rather than R packages installed from within R.
We should note, however, that only about 500 of the over
11,000 CRAN packages are available as Debian packages.

As the names `testing` and `unstable` imply, particular versions of package can
change as packages move from `unstable` into `testing`. New versions
are sent to `unstable` during the normal course of Debian development.
This can occasionally break an previously-working installation command
in a Dockerfile until the maintainer redirects the package manager to
install a package from the `unstable` sources that could previously
be installed from `testing`, or vice versa (using the `-t` option in
`apt`).  That said, packages only migrate from 
`unstable` to `testing` after a period of several days---and if the 
migration and installation of the particular version is free of 
interactions with other packages in their dependency graph. That way,
`unstable` serves as validation lab which leaves `testing` reasonably 
stable yet current.

Relative to `stable`, the `testing` stack thus offers some advantages
as almost all software can be installed through the package
manager. Installation of binary packages from `testing` generally provides
the most recent available software, and installs it quickly as a
binary. On the other hand, these Dockerfiles may require occasional 
maintenance when packages migrate and/or versions change. The resulting images 
are also inherently dynamic: rebuilding the same Dockerfile months or years 
apart will generate images with significantly different versions of software 
installed as the pool of underlying packages changes through time.
